---
title: "Chapter 2, practice"
author: "Sara"
date: "2/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.6 Practice
#Easy
2E1
Which of the expressions below correspond to the statement: the probability of rain on Monday?
(1) Pr(rain)
(2) Pr(rain|Monday) - the propability  that its raining given that its monday  - so this is true 
(3) Pr(Monday|rain)
(4) Pr(rain,Monday)/Pr(Monday) - The same as number 2 just written out. 

(2) is the correct answer. But also (4) - because its the same



2E2. Which of the following statements corresponds to the expression: Pr(Monday|rain)?
(1) The probability of rain on Monday.
(2) The probability of rain, given that it is Monday.
(3) The probability that it is Monday, given that it is raining. 
(4) The probability that it is Monday and that it is raining.

(3) is the correct answer

2E3. Which of the expressions below correspond to the statement: the probability that it is Monday,
given that it is raining?
(1) Pr(Monday|rain) - correct
(2) Pr(rain|Monday)
(3) Pr(rain|Monday)Pr(Monday)
(4) Pr(rain|Monday)Pr(Monday)/Pr(rain) - correct. Based on bayes therom, one can rewrite the probability, to the reverse.
(5) Pr(Monday|rain)Pr(rain)/Pr(Monday)


2E4. The Bayesian statistician Bruno de Finetti (1906–1985) began his book on probability theory with the declaration: “PROBABILITY DOES NOT EXIST.” The capitals appeared in the original, so I imagine de Finetti wanted us to shout this statement. What he meant is that probability is a device for describing uncertainty from the perspective of an observer with limited knowledge; it has no objective reality. Discuss the globe tossing example from the chapter, in light of this statement. What does it mean to say “the probability of water is 0.7”?
So the probability is just a random picture of how the world could look like. - and therefore not the truth og the proportion of water on the earth. So you cant just 

#Medium
2M1. Recall the globe tossing model from the chapter. Compute and plot the grid approximate posterior distribution for each of the following sets of observations. In each case, assume a uniform prior for p.
(1) W,W,W
```{r}
# define grid
p_grid <- seq(from=0 , to=1 , length.out=100)
# define prior
prior <- rep(1 , 100)
# compute likelihood at each value in grid
likelihood1 <- dbinom(3 , size=3 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior1 <- likelihood1 * prior
# standardize the posterior, so it sums to 1
posterior1 <- unstd.posterior1 / sum(unstd.posterior1)

plot(p_grid , posterior1 , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "W,W,W" )
```

(2) W,W,W,L
```{r}
# compute likelihood at each value in grid
likelihood2 <- dbinom(3 , size=4 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior2 <- likelihood2 * prior
# standardize the posterior, so it sums to 1
posterior2 <- unstd.posterior2 / sum(unstd.posterior2)

plot(p_grid , posterior2 , type="b" ,
    xlab="probability of water" , ylab="posterior probability")
mtext( "W,W,W,L")
```
(3) L,W,W,L,W,W,W
```{r}
# compute likelihood at each value in grid
likelihood3 <- dbinom(5 , size=7 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior3 <- likelihood3 * prior
# standardize the posterior, so it sums to 1
posterior3 <- unstd.posterior3 / sum(unstd.posterior3)

plot(p_grid, posterior3, type="b" ,
    xlab="probability of water" , ylab="posterior probability")
mtext("L,W,W,L,W,W,W")
```
2M2. Now assume a prior for p that is equal to zero when p < 0.5 and is a positive constant when p ≥ 0.5. Again compute and plot the grid approximate posterior distribution for each of the sets of observations in the problem just above.
```{r}
prior1 <- ifelse(p_grid < 0.5 , 0 , 1)

#the first with new prior
unstd.posterior2m21 <- likelihood1 * prior1
# standardize the posterior, so it sums to 1
posterior2m21 <- unstd.posterior2m21 / sum(unstd.posterior2m21)

plot(p_grid , posterior2m21 , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext("W,W,W, new prior")

#the second with new prior
unstd.posterior2m2_2 <- likelihood2 * prior1
# standardize the posterior, so it sums to 1
posterior2m2_2 <- unstd.posterior2m2_2 / sum(unstd.posterior2m2_2)

plot(p_grid , posterior2m2_2 , type="b" ,
    xlab="probability of water" , ylab="posterior probability")
mtext( "W,W,W,L")


#the third with new prior

# compute product of likelihood and prior
unstd.posterior2m2_3 <- likelihood3 * prior1
# standardize the posterior, so it sums to 1
posterior2m2_3 <- unstd.posterior2m2_3 / sum(unstd.posterior2m2_3)

plot(p_grid, posterior2m2_3, type="b" ,
    xlab="probability of water" , ylab="posterior probability")
mtext("L,W,W,L,W,W,W")
```
even though the plots look very weird, its okay. We are running with two different priors in the same plot. 



2M3. Suppose there are two globes, one for Earth and one for Mars. The Earth globe is 70% covered in water. The Mars globe is 100% land. Further suppose that one of these globes—you don’t know which—was tossed in the air and produced a “land” observation. Assume that each globe was equally likely to be tossed. Show that the posterior probability that the globe was the Earth, conditional on seeing “land” (Pr(Earth|land)), is 0.23.

P(Land|Earth)=1 - 0,7 = 0,3
P(Land|Mars) = 1

P(Earth) = 0.5
P(Mars) = 0.5

P(Earth|Land) = P(Land|Earth)*P(Earth)/P(Land|Earth)*P(Earth)+P(land|Mars)*P(Mars) 
```{r}
P=(0.3*0.5)/((0.3*0.5)+(1*0.5))
P
```

2M4. Suppose you have a deck with only three cards. Each card has two sides, and each side is either black or white. One card has two black sides. The second card has one black and one white side. The third card has two white sides. Now suppose all three cards are placed in a bag and shuffled. Someone reaches into the bag and pulls out a card and places it flat on a table. A black side is shown facing up, but you don’t know the color of the side facing down. Show that the probability that the other side is also black is 2/3. Use the counting method (Section 2 of the chapter) to approach this problem. This means counting up the ways that each card could produce the observed data (a black side facing up on the table).

Answer: There are 2 cards that can provide 3 blacks in different ways …. one of them is already used to give the facing card, leaving behind a 2/3 probability of the other side being black


2M5. Now suppose there are four cards: B/B, B/W, W/W, and another B/B. Again suppose a card is drawn from the bag and a black side appears face up. Again calculate the probability that the other side is black.
Answer: The same as before but now 4/5

2M6. Imagine that black ink is heavy,and so cards with black sides are heavier than cards with white sides. As a result, it’s less likely that a card with black sides is pulled from the bag. So again assume there are three cards: B/B, B/W, and W/W. After experimenting a number of times, you conclude that for every way to pull the B/B card from the bag, there are 2 ways to pull the B/W card and 3 ways to pull the W/W card. Again suppose that a card is pulled and a black side appears face up. Show that the probability the other side is black is now 0.5. Use the counting method, as before.

Answer: There are 2 ways to get a black out of 4 possibilities. Because there are two cards in the deck which is black and white. And only one card which is black/black. So the way done is the same as in the others, just that the mængde of cards are higher here do to weights. so we have 2/4 propability of getting black which is 50%

2M7. Assume again the original card problem,with a single card showing a black side face up.Before looking at the other side, we draw another card from the bag and lay it face up on the table. The face that is shown on the new card is white. Show that the probability that the first card, the one showing a black side, has black on its other side is now 0.75. Use the counting method, if you can. Hint: Treat this like the sequence of globe tosses, counting all the ways to see each observation, for each possible first card.

Answer: First we count the possibilities of having the sequence assuming the BB card is the first. This gives us 6 possible ways to produce the Black white sequence. Afterwards we assume that the BW card is the first card this only gives us 2 possible ways of producing the sequence. 6+2 =8 which tells us that all in all there are 8 ways of producing the sequence. And finding the proportion that the B/B card is the first we take 6/8 = 0.75
